---
title: "Impact of tree planting after fire on multiple forest values"
author: "I. Farnell, J. Biro, G. Greene, E. Lilles, A. Clason"
date: "2023-08-08"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    number_sections: no
    theme: flatly
    highlight: tango
    fig_caption: yes
    code_folding: hide
    df_print: paged
    keep_md: no
  word_document:
    toc: yes
    toc_depth: '2'
editor_options:
  chunk_output_type: console
---


```{r}
# Load libraries
library(data.table)
library(tidyverse)

library(treeCalcs)

files_to_source <- list.files("./R", pattern = "Function", 
                              full.names = TRUE)
sapply(files_to_source, source)

```


Pathways
```{r}

in_dir <- "./Inputs/"
out_dir <- "./Outputs/"

```


Plot treatments (planted, time since fire etc.), and all data

```{r}
FR_treatments <- fread(paste0(in_dir,"FR_Treatments.csv"))

#Plot treatment cleaning
FR_treatments[,`:=`(PlotID = as.factor(ID), Planted = as.factor(Planted))]
FR_treatments[, TimeSinceFire := 2020 - FIRE_YEAR]
#for this paper, we don't need all the columns:
plot_treatments <- FR_treatments[,.(PlotID, Planted, TimeSinceFire)]


#Tree data:
A1trees <- fread(paste0(in_dir,"A1trees.csv"))
B1trees <- fread(paste0(in_dir,"B1trees.csv"))
Regen <- fread(paste0(in_dir,"Regen.csv"))

#Soils data:
Soils <- fread(paste0(in_dir,"Soils.csv"))

#Woody debris:
cwd <- fread(paste0(in_dir,"FireRehabData_CWD.csv"),stringsAsFactors = T)
fwd <- fread(paste0(in_dir,"FireRehabData_FWD.csv"),stringsAsFactors = T)
line <- fread(paste0(in_dir,"FireRehabData_TransectDistance.csv"),stringsAsFactors = T) 

```

# Value calculations

## Basal area

```{r}
BAtrees <- BAPH(A1trees, B1trees)
PlotBAtrees <- BAtrees[,.(BAPH=sum(BAPH)), by="PlotID"]
#PlotBAtrees <- merge(PlotBAtrees, FR_treatments[,.(PlotID, Planted, TimeSinceFire)])

```


## Basal area by diameter class
no done this one - need to go long to wide (so the columns represent diam bins BA)
```{r}
# 7.5+, 12.5+, 17.5+, 22.5+ (size classes from BC VRI )
BAdiamClass <- BAPHsizeCl(A1trees, B1trees)
PlotBAdiam <- BAdiamClass[,]
BAdiamClass <- merge(BAdiamClass, FR_treatments)


```

## Stems per hectare

```{r}
#live trees per hectare ------------------
PlotTree <- TreeDensity(A1trees, B1trees) # ask Alana why 2cm size bins
# AC - changed it to 5cm (2 is for SORTIE)
# get live tree size classes across species
PlotTree_allSp <- PlotTree[, .(SPH = sum(SPH)), by = c("PlotID","DBH_bin")]
PlotTree_allSp_d <- dcast(PlotTree_allSp[,.(PlotID, DBH_bin, SPH)], 
                        PlotID ~ DBH_bin,
                        value.var = "SPH", fun = sum)
#merge with treatment data
PlotTree_tr <- merge(PlotTree_allSp_d, 
                     FR_treatments[,.(PlotID, Planted, TimeSinceFire)], by = "PlotID",
                     all.y = TRUE)
PlotTree_tr[, Planted := as.factor(Planted)]

cols_to_replace <- c("5","10","15","20","25","30","35","40")
PlotTree_tr[, (cols_to_replace) := lapply(.SD, function(x) ifelse(is.na(x), 0, x)), 
                 .SDcols = cols_to_replace]
PlotTree_tr


# Snags per hectare ----------------------
#not done
PlotSnags <- SnagDensity(A1trees, B1trees)
PlotSnags <- merge(PlotSnags, FR_treatments[,.(PlotID, Planted, TimeSinceFire)])


```

## CWD and cwd habitat

```{r}
# CWD HQI
PlotHQI <- cwdHQI(cwd)
PlotCWD <- cwdVar(cwd)
PlotCWD <- merge(PlotCWD, PlotHQI)

# CWD volume
CWDvol <- cwdVol(cwd, line) #volume by species & decay class
PlotCWDvol <- CWDvol[,.(VolHa=sum(VolHa)), by="PlotID"]
PlotCWD <- merge(PlotCWD, PlotCWDvol)
#PlotCWD <- merge(PlotCWD, FR_treatments[,.(PlotID, Planted, TimeSinceFire)])

```

## Merchantable Volume:

```{r}
#MerchVolCalcConst <- fread("./Inputs/MerchVolCalcConst.csv")
merchTrees <- B1trees

t <- vector()
for(i in 1:nrow(merchTrees)){
  t[i] <- TreeMerchVolFN(Species=merchTrees[i,Species], DBH=merchTrees[i,DBH], 
                       HT=merchTrees[i,Height])
}

merchTrees[, MerchVol:=t]
merchTrees[is.na(MerchVol), MerchVol:= 0] #check the function to make sure the NAs are correct

PlotmerchTrees <- merchTrees[, .(MerchVol = sum(MerchVol)), by = "PlotID"]

PlotmerchTrees
```

## Wildfire:
```{r}

regen_biomass <- calc_regen_biomass(Regen = Regen)
fwrite(regen_biomass, "./Inputs/regen_biomass.csv")

```




## Forest carbon:

need to check all these values line up with our paper. Made slight changes to code!
Also - need to get unknown conifer and unknown cwd estimates (Ingrid review treeCalcs)
```{r}
# trees ------------------
treeC <- calc_tree_carbon(A1trees = A1trees, B1trees = B1trees)

regenC <- calc_regen_c(Regen = Regen)

carbon_dat <- merge(plot_treatments, merge(treeC, regenC, by = "PlotID", all = TRUE),
                   by = "PlotID", all = TRUE)
for (i in names(carbon_dat))carbon_dat[is.na(get(i)), (i):=0]

# roots -------------------
carbon_dat[, LiveRootC := ((LiveMgHa * 0.222) + (Regen_L_MGHa * 0.222)), 
          by = seq_len(nrow(carbon_dat))] #root =  0.222 x live biomass
#decaying the tree roots by time since fire
carbon_dat[, DeadRootC := ((DeadMgHa * exp(-0.1044 * TimeSinceFire)) + 
                             (Regen_D_MGHa * 0.222)), 
          by = seq_len(nrow(carbon_dat))] #root =  0.222 x  dead biomass
  
# soil --------------------
soilC <- calc_soil_carbon(soil = Soils)

carbon_dat <- merge(carbon_dat, soilC, by = "PlotID", all = TRUE)
for (i in names(carbon_dat))carbon_dat[is.na(get(i)), (i):=0]

# woody debris ------------
# need to double check mean values for UC in SBS
wdC <- calc_wd_carbon(cwdDat = cwd, fwdDat = fwd, lineDat = line)

carbon_dat <- merge(carbon_dat, wdC, by = "PlotID", all = TRUE)
for (i in names(carbon_dat))carbon_dat[is.na(get(i)), (i):=0]
 

# all together ------------------------------
carbon_dat_m <- melt(carbon_dat, 
                    id.vars = c("PlotID","Planted","TimeSinceFire"),
                    measure.vars = c("LiveMgHa","DeadMgHa","Regen_D_MGHa","Regen_L_MGHa",
                                     "MinSoilC_Mgha","Litter_C_MgHa",
                                     "ForestFl_MgHa","cwd_MgHa","fwd_MgHa","LiveRootC",
                                     "DeadRootC"),
                    variable.name = "CarbonSource",
                    value.name = "CarbonMgHa")

carbon_dat <- merge(carbon_dat[,.(PlotID, Planted,TimeSinceFire)],
                    carbon_dat_m[,.(TotalCarbon = sum(na.omit(CarbonMgHa))),
                                             by="PlotID"], by ="PlotID")

carbon_dat <- merge(carbon_dat,carbon_dat_m[CarbonSource=="DeadMgHa"|
                                                   CarbonSource=="Regen_D_MGHa"|
                                                   CarbonSource=="cwd_MgHa"|
                                                   CarbonSource=="fwd_MgHa",
                                                 .(DeadCarbon = sum(na.omit(CarbonMgHa))),
                                                 by="PlotID"],by="PlotID")

carbon_dat <- merge(carbon_dat,carbon_dat_m[CarbonSource=="LiveMgHa"|
                                                   CarbonSource=="Regen_L_MGHa"|
                                                   CarbonSource=="LiveRootC",
                                                 .(LiveCarbon = sum(na.omit(CarbonMgHa))),
                                                 by="PlotID"],by="PlotID")

carbon_dat

```


## Moose habitat

```{r}

```

## OTHER BIODIVERSITY????

```{r}

```


# Combined values
need to add a bunch more here when done
```{r}
value_dat1 <- merge(PlotBAtrees, PlotTree_tr,by = "PlotID", all.y = TRUE)
value_dat1 <- merge(value_dat1, PlotCWD, by = "PlotID")
value_dat1 <- merge(value_dat1, PlotmerchTrees, by = "PlotID")
value_dat <- merge(value_dat1, carbon_dat[,.(PlotID,TotalCarbon,DeadCarbon)], by = "PlotID")
                    
```


prelim figures
```{r}
# Basal area
BA <- ggplot(value_dat, aes(x=TimeSinceFire, y= BAPH)) +
  geom_point(aes(colour=Planted)) +
  geom_smooth(method="lm",aes(colour=Planted,fill=Planted))
BA

# Basal area diameter classes - this one won't work yet
BAcl <- ggplot(value_dat, aes(x=DBH_bin, y=BAPH)) +
  geom_col(aes(colour=Species), position="dodge") +
  facet_grid(rows=vars(Planted))
BAcl

# large live trees - this one won't work yet
lglive <- PlotTree %>%
  filter(DBH_bin >=20) %>%
  ggplot(aes(x=DBH_bin, y=SPH)) +
  geom_point(aes(colour=Planted)) +
  facet_grid(rows=vars(Planted))
lglive

# large cottonwood or aspen - this one won't work yet
liveAcAt <- PlotTree %>%
  filter(Species=="Ac" | Species=="At") %>%
  filter(DBH_bin >= 20) %>%
  ggplot(aes(x=DBH_bin, y=SPH)) +
  geom_point(aes(colour=Species)) +
  facet_grid(rows=vars(Planted))
liveAcAt

# large snags
snags <- PlotSnags %>% 
  filter(DBH_bin >= 20) %>%
  ggplot(aes(x=DBH_bin, y=SPH)) +
  geom_point(aes(colour=Planted)) +
  facet_grid(rows=vars(Planted))
snags

# CWD
hqi <- ggplot(PlotCWD, aes(x=Planted, y=HQI)) +
  geom_boxplot()
hqi

cwdVol <- ggplot(PlotCWD, aes(x=Planted, y=VolHa)) +
  geom_boxplot()
cwdVol

cwdMaxDiam <- ggplot(PlotCWD, aes(x=Planted, y=diamMax)) +
  geom_boxplot()
cwdMaxDiam

cwdN <- ggplot(PlotCWD, aes(x=Planted, y=PlotN)) +
  geom_boxplot()
cwdN

cwdLoc <- cwd %>%
  mutate(Location_m = if_else(condition = Transect==2,
                              true=Location_m+50,
                              false=Location_m)) %>% 
  ggplot(aes(x=Location_m, y=PlotID)) +
  geom_point() +
  facet_grid(rows=vars(Planted))
cwdLoc # i am able to map cwd piece location - not sure how to put this into piles & dispersed


```










